---
title: "Introduction to R, Part 2"
author: "CSE Training"
date: ""
output: html_document
---

We are going to do another demo project to demonstrate how to do some advanced data analysis using R. Specifically, we will cover topics including regression models, hypothesis tests and reproducible research. 

We will use a built-in data set of R, **mtcars**, to do this project. The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973 to 1974 models). We are interested in the following questions:

1. Is there a difference in fuel efficiency between cars with automatic and manual transmissions? 

2. Can we develop a regression model for fuel efficiency using all other variables in the data set?

## Read Data and Display Data

```{r import}
# Read data and take a quick look at it
data(mtcars);
str(mtcars);
```

## Difference between automatic and manual transmissions

First, plot some exploratory graphs. 

```{r explore, fig.width=10, fig.show='asis'}
# Extract the data we need
automatic = mtcars[mtcars$am == 0, ]$mpg; 
manual = mtcars[mtcars$am == 1, ]$mpg;

# Plot the graphs
par(mfrow = c(1,2));
plot(automatic, col="red", pch=18, main = "Figure 1", xlab = "", ylab = "Miles per Gallon");
points(manual, col="green", pch=18);
legend("topright",c("Automatic", "Manual"), pch = c(18, 18), col = c("red", "green"));

boxplot(mpg~am,data=mtcars,main="Figure 2",varwidth=TRUE, 
        col=c("green","red"), names=c("Automatic","Manual"));
```

From the graph, we see that most points of manual transmission are above those of automatic transmission. In order to see the difference statistically, we perform a [student's t-test](http://en.wikipedia.org/wiki/Student's_t-test) about the null hypothesis that automatic and manual transmission have the same mean value for miles per gallon(mpg). 

```{r ttest}
# Do the t-test for difference between means
t.test(automatic, manual, alternative="two.sided", conf.level=0.99);
```

So, with a confidence level of 99%, we can claim that in 1973-74 models, automatic transmissions have smaller mpg values than manual transmissions. 

## Regression Model

In the previous section, we see that the variable **mpg** varies much for different values of the variable **am**. In this way, we will definitely have the variable **am** inside the regression model for **mpg**. Further more, we would also like to include interaction terms of **am** with other variables. So, what we will do is: include all variables inside the model (both variables and interaction terms with am). Then, we will remove variables from the model gradually to get our final model.

```{r models_appen}
# Include all variables and interaction terms with am
fit <- lm(mpg~as.factor(am) *., data = mtcars);
summary(fit);
```

Then we choose variables that are significant in the model, which are **as.factor(am)**, **as.factor(am):wt**, **as.factor(am):qsec** and **as.factor(am):carb** and do a nested model test. 

```{r nested}
fit1 <- lm(mpg ~ as.factor(am), data = mtcars);
fit2 <- update(fit, mpg~ as.factor(am):wt);
fit3 <- update(fit, mpg~ as.factor(am):(wt + qsec));
fit4 <- update(fit, mpg~ as.factor(am):(wt + qsec + carb));
anova(fit1, fit2, fit3, fit4);
```

From the result, fit3 with formula: **mpg ~ as.factor(am):(wt + qsec)** is the best model among these four models. We do the regression and see the results for coefficients and residuals. 

```{r tempmod}
summary(fit3);
```

Compared the adjusted R squared between fit3 and fit (0.879 and 0.903), this could be the model we use to do analysis between **mpg** and **am**. A plot for residuals is shown in Figure 3. From the plot we can see that the model fits well with few outliers. 

```{r residualplot}
# Plot residuals of the final model.
par(mfrow = c(2,2), oma = c(0,2,4,2))
plot(fit3);
mtext("Figure 3 Residual Diagnostics", cex = 1.5, outer = TRUE, line = 1)
```

**Analysis of the model**: If we hold the **qsec** constant at its mean, for each unit increase in the weight starting at mean, miles per gallon will decrease -3.176 for manual cars and -6.1 for automatic cars. These coefficients are quite significant in the model with low standard errors. On the other hand, if we hold the weight constant at its mean, for each unit increase in the qsec starting at mean value, manual cars will make 0.834 more miles per gallon while 1.446 for automatic cars.


## Reproducible research

The R package, *knitr*, provides a great tool to make your research reproducible. It allows you to include R code, graphs and many things else inside a markdown file. Please go to [KnitR](http://yihui.name/knitr/): http://yihui.name/knitr/ to find more help files. 




